---
title: "hw2"
author: "Audrey Comer"
date: "1/31/2019"
output: html_document
---

```{r}
a0 <- 1
a0
```
1. Today's temperature has a high of 15 and a low of -10 degree fahrenheit.  Find the corresponding degrees in celsius and storage it in a (single) vector named `cel`. Print the variable.
```{r}
f <- c(-10:15)
cel <- (5/9*f)-32
cel
```
2. Create a vector of length 200. The odd entries should be 1 and even entries should be 0. Storage the created in a variable named `vec200` and print.
```{r}
vec200 <- rep(c(1,0),100)
vec200
```
3. How many perfect squares are there between 1 and 1001? Storage the answer in a  variable named `numPerfect` and print. You may want to look at `?round`.
```{r}
sqrt(1001)
numPerfect <- floor(sqrt(1001))
numPerfect
```
```{r}
library(ggplot2)
```
What are the 3 most fuel-efficient models in terms of highway mileage? Storage the data frame (with 3 rows) containing all information of these three cars in a variable named top3 and print.
```{r}
summary(mpg)
highway <- mpg$hwy
highway
names(highway) <- mpg$model
highway
sort(highway, decreasing = TRUE)
head(highway, 3)
top3model <- names(head(highway,3))
top3model

top3 <- mpg[mpg$hwy >40, ]
top3
```

How many different compact models are included in this dataset? Storage the answer in a variable named numCompact and print.
```{r}
mpg$class
compact <- mpg[mpg$class == "compact", ]
compact
numCompact <- nrow(compact)
numCompact
```

Create a scatterplot of hwy and cty and describe the relationship. Why are there so few points visible?
So few points are visible because the data overlaps -- there are multiple data points with the same (x,y) (hwy,cty) values as each other.
```{r}
plot(mpg$hwy, mpg$cty)
```



Are cars in 2008 more fuel efficient than cars in 1999? Use graphs and summary statistics to answer this question.
```{r}
y1999 <- mpg[mpg$year==1999,] #creates df of just 1999
y1999
y2008 <- mpg[mpg$year==2008,] #creates df of just 2008
y2008
y1999$mpg_avg <- rowMeans(y1999[c('hwy', 'cty')], na.rm=TRUE) #adds row for mean of hwy and city
y2008$mpg_avg <- rowMeans(y2008[c('hwy', 'cty')], na.rm=TRUE) #ditto
df <- rbind(y1999,y2008) #creates new df combining 1999 and 2008
```
Below is a box plot that show the average for 1999 was higher than 2008. However, outliers factor into this. Cars from 2008 had a wider distribution of MPG, especially for cars above the average.
```{r}
?boxplot
boxplot(mpg_avg~year,data=df, col=c('red','blue'), main='Average MPG by Year',xlab='Year',ylab='MPG')
```

The third diagram below is a histogram showing average mpg of 1999 in red and 2008 in blue. We can see again that 2008 (blue) has a wider distribution in the middle range. 
```{r}
p1 <- hist(x=y1999$mpg_avg,breaks=12,xlim=c(10,40)) #setting up 1999 histogram
p2 <- hist(x=y2008$mpg_avg,breaks=12,xlim=c(10,40)) #setting up 2008 histogram
plot(p1,col=rgb(1,0,0,1/4),xlab="Miles per gallon",ylab="Frequency",main="Distribution of MPG for cars from 1999 (red) and 2008 (blue)",xlim=c(10,40))
plot(p2,col=rgb(0,0,1,1/4),xlim=c(10,40),add = TRUE)
```
```{r}
summary(y1999)
```
```{r}
summary(y2008)
```
The mean mpg value (averages) for 1999 are higher in the city and averaged categories. The mean highway mpg is higher in 2008 by .02 mpg. Overall, miles per gallon was higher on average in 1999. 
